\bf Model Visualization Techniques for a Social Network Model

### 2018-09-04 paper-child.Rnw
committing just in case

The second method of model visualization we use is visualizing collections of models. There are many possible ways to collect SAOMs together, so %. We could look at the same model fit to different data, different models fit to the same data, or we could fit the same models to the same data many times to see how the simulations change. F
we chose four ways to create collections of models that would give us the most insight: (1) exploring the space of all possible models, (2) fitting the same model to the same data many times, (3) varying model settings, and (4) fitting the same model form to different data. We chose these four collections because they each explore something different about SAOMs. The first considers the many parameters to include in a SAOM, while the second results in different parameter estimates, so we can see how parameter estimates vary. The third collection shows how those many possible parameters affect the values of parameters in the fitted models. Finally, we explore how the same model behaves under two completely different circumstances.

\subsection{Exploring the space of all possible models}\label{sec:allpossible}

The \texttt{RSiena} manual contains over eighty effects to include in a SAOM's objective function. In order to select parameters to include in the models for our example data, we searched through the possible effects available to model given the data structure to find \textit{significant} effects. We tested for significance using the Wald-type tests built into \texttt{RSiena} for one-at-a-time effects testing. We start with the outdegree and reciprocity measures as the foundation of the models we fit, then add one effect, fit the model, test the additional effect for significance, and repeat for all possible parameters to add to the model. We performed this procedure, which we used to pick M2 and M3, for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in Figure~\ref{fig:sigtesting} in the Appendix. %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

\subsection{Fitting the same model to the same data}\label{sec:samemodsamedat}

@@

After fitting models M1, M2, and M3 to the friendship network 1,000 times each, we examine the distribution of the fitted values, which are shown in Figure~\ref{fig:distests}. We can see from these distributions that the inclusion of the JTT parameter, $\beta_3$ is obviously affecting the distributions of the other four parameters included in all models, $\alpha_1$, $\alpha_2$, $\beta_1$, and $\beta_2$, while these parameters are less affected by the inclusion of $\beta_4$. When $\beta_3$ is included, its estimate is postitive, meaning that friendships between two girls with different drinking behaviors tend to form when there is an intermediary who is already friends with the two girls. The inclusion of this parameter leads to increases in the rate parameters' estimates, suggesting that the presence of the transitive triplet behavior means that the girls would also change friends more frequently. It is not clear, however, that the addition of a parameter to the objective function \textit{should} effect the estimation of the rate parameters. The strong correlations between the estimates of the rate parameters and of the objective function parameters are shown in Figure~\ref{fig:corplots} in the Appendix. In addition, the outegree parameter, $\beta_1$ decreases when $\beta_3$ is included, while the reciprocity parameter, $\beta_2$ increases. This implies the girls in the data prefer to form small, closely knit friend groups, as indicated by reciprocated ties and jumping transitive triplet formation, as opposed to being popular and having many friends. Thus, having many friends who do not reciprocate is more stronlgy discouraged by M2 than in the other models. %In comparison with models M1 and M3, model M2 typically has higher estimates of the rate parameter, meaning that the inclusion of the covariate statistic in the model leads to higher estimates of the number of times, on average, a node gets to change its ties.

\subsection{Varying model settings}

The next method of viewing collections of models is with varying model settings. We have varied model settings already by choosing three different models M1, M2, M3 to fit to our example data sets. In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examined the parameter densities. Here, we simulate from these three models. For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model. We simulated 1,000 observations from each of the three models in this way. From these simulations, we created visualizations that represent an average network from models M1, M2, and M3 using the same method described in Section~\ref{sec:minds}. These averages are shown in Figure~\ref{fig:averagenet} in the Appendix. We see that all models are failing to capture the structure of the second wave.

\subsection{Fitting the same model to different data}\label{sec:samemoddiffdat}

In addition to fitting M1, M2, and M3 to the friendship data 1,000 times, we also fit them to the senate data 100 times. The means and standard deviations of the parameter estimates for each combination of model and data are given in Table~\ref{tab:meanests}, while the density plots of each parameter in the model for each data set is given in Figure~\ref{fig:compareModels}.

>>>>

In addition to assessing model fit, we are also interested in model behavior so that we can choose better fitting models in the future. We examine model behavior by visualizing collections of models. There are many possible ways to collect SAOMs together, so 
we chose four ways to create collections of models that would give us the most insight: (1) exploring the space of all possible models, (2) fitting the same model to the same data many times, (3) varying model settings, and (4) fitting the same model form to different data. We chose these four collections because they each explore something different about SAOMs. The first considers the many parameters to include in a SAOM, while the second results in different parameter estimates, so we can see how parameter estimates vary. The third collection shows how those many possible parameters affect the values of parameters in the fitted models. Finally, we explore how the same model behaves under two completely different circumstances.

%\subsection{Exploring the space of all possible models}\label{sec:allpossible}

\noindent \textbf{Exploring the space of all possible models:} The \texttt{RSiena} manual contains over eighty effects to include in a SAOM's objective function. In order to select parameters to include in the models for our example data, we searched through the possible effects available to model given the data structure to find \textit{significant} effects. We tested for significance using the Wald-type tests built into \texttt{RSiena} for one-at-a-time effects testing. We start with the outdegree and reciprocity measures as the foundation of the models we fit, then add one effect, fit the model, test the additional effect for significance, and repeat for all possible parameters to add to the model. We performed this procedure, which we used to pick M2 and M3, for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in Figure~\ref{fig:sigtesting} in the Appendix. %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

%\subsection{Varying model settings}

\noindent \textbf{Varying model settings:} The next method of viewing collections of models is with varying model settings. We have varied model settings already by choosing three different models M1, M2, M3 to fit to our example data sets. In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examined the parameter densities. Here, we simulate from these three models. For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model. We simulated 1,000 observations from each of the three models in this way. From these simulations, we created visualizations that represent an average network from models M1, M2, and M3 using the same method described in Section~\ref{sec:minds}. These averages are shown in Figure~\ref{fig:averagenet} in the Appendix. We see that all models are failing to capture the structure of the second wave.

%\subsection{Fitting the same model to the same data}\label{sec:samemodsamedat}

@@

\noindent \textbf{Fitting the same model to the same data:} After fitting models M1, M2, and M3 to the friendship network 1,000 times each, we examine the distribution of the fitted values, which are shown in Figure~\ref{fig:distests}. We can see from these distributions that the inclusion of the JTT parameter, $\beta_3$ is obviously affecting the distributions of the other four parameters included in all models, $\alpha_1$, $\alpha_2$, $\beta_1$, and $\beta_2$, while these parameters are less affected by the inclusion of $\beta_4$. When $\beta_3$ is included, its estimate is postitive, meaning that friendships between two girls with different drinking behaviors tend to form when there is an intermediary who is already friends with the two girls. The inclusion of this parameter leads to increases in the rate parameters' estimates, suggesting that the presence of the transitive triplet behavior means that the girls would also change friends more frequently. It is not clear, however, that the addition of a parameter to the objective function \textit{should} effect the estimation of the rate parameters. The strong correlations between the estimates of the rate parameters and of the objective function parameters are shown in Figure~\ref{fig:corplots} in the Appendix. In addition, the outegree parameter, $\beta_1$ decreases when $\beta_3$ is included, while the reciprocity parameter, $\beta_2$ increases. This implies the girls in the data prefer to form small, closely knit friend groups, as indicated by reciprocated ties and jumping transitive triplet formation, as opposed to being popular and having many friends. Thus, having many friends who do not reciprocate is more stronlgy discouraged by M2 than in the other models. %In comparison with models M1 and M3, model M2 typically has higher estimates of the rate parameter, meaning that the inclusion of the covariate statistic in the model leads to higher estimates of the number of times, on average, a node gets to change its ties.

%\subsection{Fitting the same model to different data}\label{sec:samemoddiffdat}

\noindent \textbf{Fitting the same model to different data:} In addition to fitting M1, M2, and M3 to the friendship data 1,000 times, we also fit them to the senate data 100 times. The means and standard deviations of the parameter estimates for each combination of model and data are given in Table~\ref{tab:meanests}, while the density plots of each parameter in the model for each data set is given in Figure~\ref{fig:compareModels}.
